---
title: "Graphical LASSO support"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graphical LASSO support}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(CARlasso)
```

We also have Graphical LASSO support for study with no predictors, the main interface is `bGlasso`, please also see the [reference](https://yunyishen.ml/CAR-LASSO/dev/reference/bGlasso.html).

## Network for Gaussian data

This is the case when data is Gaussian or can be transformed to Gaussian. We only need responses in this case

```{r ar1data}
set.seed(42)
dt <- simu_AR1(n=100,k=5, rho=0.7)
dt <- dt[,1:5]
head(dt)

```
Different from `CARlasso` we do not really need to have a formula. To use the Normal version, we should set `link="identity"` which is the default. 


```{r ar1example_first}
glassores <- bGlasso(data = dt)
plot(glassores)
```

We could see the false link due to the common input.

## Network for compositional data

This is common in microbe related studies and some ecological applications with relative abundance only. For instance, microbe relative abundance came from sequencing and in this case we will have the sum of "abundance" decided by the sequence depth rather than the real total abundance. The data is usually described as "compositional". In `CARlasso` this type of data were modelled as Logit-Normal-multinomial. In this case, we have less relative abundance need to have a "reference level" taxa and all others are "relative" to this taxa, we will give a reduced version:

Have a look of the data, still a dataframe with all predictors and responses

```{r comp_data}
dt <- mgp154[,c("Alistipes","Bacteroides",
                        "Eubacterium","Parabacteroides","all_others")]
```

To run the composition model, we need to set `link="logit"`

```{r compositional1}
gut_res <- bGlasso( data = dt,link = "logit", 
                    n_iter = 2000, 
                    n_burn_in = 1000, thin_by = 2)
plot(gut_res)
```

Note that in this case the last one in the dataframe will be the reference level 

## Network for counting data

This is common in a lot of ecological applications. For instance number of seedlings within a site. The responses were count rather than continuous. In `CARlasso` it is modelled as Poisson with rate follow log-Normal. We will give the example using the data above since the responses are indeed count, but keep in mind we need to be careful whether a couting data is real counting or compositional. To distinguish them, ask the question: **Is the sum know/decided by us/not free**? If yes, we want to use compositional models. 

To run the composition model, we need to set `link="log"`

```{r counting}
gut_res <- gut_res <- bGlasso( data = dt[,1:4],link = "log", 
                    n_iter = 2000, 
                    n_burn_in = 1000, thin_by = 2)
plot(gut_res)
```



